---
title: "SLDPCA and Residential Markets"
subtitle: 'King County, WA Analysis'
author: "Andy Krause"
date: "12/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = "html") 
```

```{r, echo=FALSE, }
   setwd('..')
   # Read in Data
   summary_df <- readRDS(file.path(getwd(), paste0('summary.rds')))
   int_df <- readRDS(file.path(getwd(), paste0('interpretability.rds')))

```
## What I Did

* Ran a host of models on King County home sales from 2017 to end of 2019
    * Linear and Random Forest Models
    * Variety of different specifications aimed at teasing out the marginal contribution of the SLD variables
* Extracted measures of impact from each variable in each model
    * Linear = coefficient values
    * Random Forest = Partial Dependency analysis on each SLD variable
* Summarized model performance
    * Model Fit (adj R-squared)
    * Predictive Ability: Median Absolute Percent Error (MdAPE)
            * Out-of-time prediction on last month in the data
    * Spatial Autocorrelation of Errors (SPAC): Moran's I Statistic, lower is better
    * Predictive Ability of Sales Near Submarket Boundary (Border MdAPE)
    
## Interpreting the SLD Impact

The plot below shows the impact (in % terms) of each of the SLD variables in four different model specifications:

* Linear (Blue) vs Random Forest (Orange)
* SLD as only spatial variable (Solid Line) vs SLD with X,Y and Submarket Fixed Effects (Dashed)

Broad takeaways:

* SLDs 1, 2 and 4 are positive, 3 and 5 negative
* The magitude of the impacts are greatly reduced when other spatial control features are added, but still remain (other than SLD 4)
* Linear and Random Forest models show similar directionality in all cases and roughly similar magnitude for all SLD variables except SLD 2 where the linear model suggests a much larger impact than the random forest models.
* SLDs 1, 3 and 5 show an 'S' curve in the random forest impacts suggesting that even with the standardization the linear model is missing some variation across the spectrum of SLD variable values. 


```{r, echo = FALSE}

 int_df <- int_df %>%
  dplyr::mutate(sld_var = gsub('sld', 'SLD ', sld_var))

 ggplot(int_df, aes(x = var_stds, y= value, color = model, linetype = spec)) +
    facet_wrap(~sld_var) +
    geom_line() +
    scale_color_manual(name = 'Model Type',
                       values = c('blue', 'orange'),
                       labels = c('Linear', 'Random Forest')) +
    scale_linetype_manual(name = 'Model Specification',
                          values = c(2,1),
                          labels = c('All Spatial', 'SLD Only')) +
    xlab('\nSLD Variable Value (Normalized)') +
    ylab('Impact on Value (%)\n') +
    theme(legend.position = 'bottom')

```


```{r, echo = FALSE, warning = FALSE}

  summary_df <- summary_df %>%
  dplyr::mutate(fit = round(fit, 3),
                spac = round(spac, 1),
                border_mdape = round(border_mdape, 3))
 
  summary_tbl <- summary_df %>%
    dplyr::select(-model) %>%
    knitr::kable(.) %>%
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover"), full_width = F) %>%
    kableExtra::pack_rows("Linear Model", 1, 7) %>%
    kableExtra::pack_rows("Random Forest", 8, 14)


```

### Model Performance

Next, I looked at how well the SLD features helped a variety of model performance indicators. 

Key takeaways:

* Model fit and predictive abilty are highly correlated, will just talk about predictions
* SLD variables don't add much to model performance in linear model
* Impacts of SLD variables on model performance much more significant in random forest model
* Submarket fixed effects are hard to beat in a linear model
* Submarket fixed effects are much less effective than X,Y coordinates and/or SLD variables in random forest
* All model show significant spatial autocorrelation
* Analysis of Border properties is inconclusive...needs a bit more work here. 

```{r, echo=FALSE}
  summary_tbl
```


